### 2024-02-08 docs/manuscript/manuscript.qmd
Referencias añadidas introducción y mayteriales y metodos

The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being. Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process. In recent times, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.

@@

Cardiac autonomic response to physical exercise was measured as previously validated for this age group (Castillo-Aguilar et al., 2023). The protocol is briefly described below:

HRV was recorded through a non-invasive chest band, immediately before, during, and immediatly after the execution of the two-minutes step test (TMST). Vital signs, including blood pressure, were monitored throughout the test, and participants' well-being was visually checked to ensure they were comfortable and mentally prepared. For resting HRV measurements (prior and posterior to TMST), the volunteers remained seated in a chair, while proper support of the feet and back, and avoiding talking during the recording was ensured. RR intervals were recorded continuously during the last 10 min of rest, and were subsequently analyzed for 5 min in each occasion. The breathing rate was spontaneous. As part of the protocol, it was ensured that the participant had a blood pressure less than 140/90 mmHg to start the HRV measurements.

@@

SPPB consists of a physical test used to measure three components of physical fitness, described below (Guralnik et al., 1994):

@@

Body mass (kg) and total body fat (%) were assessed by bioimpedance using the Tanita BC-558 Ironman Segmental Body Composition Monitor (Tanita Ironman, Arlington Heights, IL 60005, USA), with a concordance of 89.3% compared to the Dual X-ray Absorption test using standard measurement protocols (Mialich et al., 2011).

@@

The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity (Rikli, 1999, Castillo-Aguilar et al., 2023). The test consists of a two-minute assessment designed to evaluate cardiorrespiratory fitness. Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.

>>>>

The process of aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being[@pmid:30185958], [@pmid:31154942]. Body composition and cardiovascular function are particularly relevant, as they are closely associated with the aging process [@pmid:36948782], [@pmid:36443946], [@pmid:30065258], [@pmid:36918543], [@pmid:32483625]. In recent times, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.

@@

Cardiac autonomic response to physical exercise was measured as previously validated for this age group [@ijerph20054456]. The protocol is briefly described below:

HRV was recorded through a non-invasive chest band, immediately before, during, and immediately after the execution of the two-minutes step test (TMST). Vital signs, including blood pressure, were monitored throughout the test, and participants' well-being was visually checked to ensure they were comfortable and mentally prepared. For resting HRV measurements (prior and posterior to TMST), the volunteers remained seated in a chair, while proper support of the feet and back, and avoiding talking during the recording was ensured. RR intervals were recorded continuously during the last 10 min of rest, and were subsequently analyzed for 5 min in each occasion. The breathing rate was spontaneous. As part of the protocol, it was ensured that the participant had a blood pressure less than 140/90 mmHg to start the HRV measurements.

@@

SPPB consists of a physical test used to measure three components of physical fitness, described below [@Guralnik1994]:

@@

Body mass (kg) and total body fat (%) were assessed by bioimpedance using the Tanita BC-558 Ironman Segmental Body Composition Monitor (Tanita Ironman, Arlington Heights, IL 60005, USA), with a concordance of 89.3% compared to the Dual X-ray Absorption test using standard measurement protocols [@Mialich2011].

@@

The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999], [@ijerph20054456]. The test consists of a two-minute assessment designed to evaluate cardiorrespiratory fitness. Participants are instructed to raise their right knee to a marked point on a wall as many times as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint.
